---
title: "Download demographic predictors and ZCTA shapefile"
author: 
  - "Cole Tanigawa-Lau"
date: ""
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float: yes
  theme: paper
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r about, echo=FALSE, results = "asis"}
cat(
  sprintf("Updated: %s.", 
          format(Sys.time(), "%b %d, %Y at %H:%M:%S", usetz = TRUE)),
  sprintf("Working directory: `%s`.", getwd()),
  sep = "\n\n"
)
```


The code chunks were run interactively and are not evaluated when knitting the document.

Reads:

- None

Writes:

 - data/acs/*
 - data/acs/alltracts17.rds
 - data/acs/alltracts10-15_lng.rds
 - data/census00_lng.rds

# Setup

```{r packages, warning=FALSE, message=FALSE}
rm(list = ls())


library(parallel)
library(coler)
library(data.table)
library(tigris)
library(tidycensus)
```

```{r}
# apis <- listCensusApis()
# 
# acsvars <- listCensusMetadata(name = "acs/acs5", vintage = "2009", type = "variables")
# View(acsvars)
```

# Choose variables
The 5-year ACS has nationwide coverage but starts in 2009.


2018 variable codes: https://api.census.gov/data/2018/acs/acs5/variables.html

2009 variable codes: https://api.census.gov/data/2009/acs/acs5/variables.html
```
B00001_001 # population count


# nino: B26101_009  # 65-74
# nino: B26101_010  # 74-85
# nino: B26101_011  # 85+

# from load_variables(2011, "acs5)
B06001_011  # 65-74
B06001_012  # 74-85

B99163_004 # ability to speak english in population age 5+
B10054_013  # speak english less than "very well"

B02001_002  # total white alone
B02001_003  # total black alone

B28001_001 # number of computers in household
B28001_011 # no computer
B28002_013 # no internet access

B06009_003 # total hs grads
B06009_002 # total less than hs

B07101_002 # non-movers

B06011_001 # median income in past 12 months
B19313_001 # aggregate income in past 12 months

B16009_001 # total poverty status in past 12 months
B06012_002 # total below 100% of poverty level
B06012_003 # total between 100 and 149% of poverty level
B10059_001 # total poverty status in past 12 months of childrearing grandparents
B19057_001 # public assistance income in past 12 months
B19055_001 # social security income in past 12 months
```


```{r}
vars <- 
  c("B00001_001", # population count
    
    "B06001_011",  # 65-74
    "B06001_012",  # 74-85
    "B06001_013",  # 86+
    
    "B99163_004",  # ability to speak english in population age 5+ (imputed)
    "B10054_013",  # speak english less than "very well"
    
    "B02001_002",  # total white alone
    "B02001_003",  # total black alone
    
    "B06009_003", # total hs grads
    "B06009_002", # total less than hs
    
    "B07101_002", # non-movers
    
    "B06011_001", # median income in past 12 months
    "B19313_001", # aggregate income in past 12 months
    
    "B16009_001", # total poverty status in past 12 months
    "B06012_002", # total below 100% of poverty level
    "B06012_003", # total between 100 and 149% of poverty level
    "B10059_001", # total poverty status in past 12 months of childrearing grandparents
    "B19057_001", # public assistance income in past 12 months
    "B19055_001"  # social security income in past 12 months
  )

compvars <- c("B28001_001", "B28001_011", "B28002_013")
```

```{r}
data(fips_codes)
setDT(fips_codes)

uniq_state <- unique(fips_codes$state)

fips_list <- split(fips_codes, uniq_state)
```

```{r}
# Functions to download data on all census tracts for a given state-year.
# Rather than return the data set, these functions save data to disk.
# Not the most efficient, but good for monitoring progress and recovering from a failed download.


## ACS 5-year
acs_state <- function(state, county = NULL, geography, variables, year, geometry, ...) {
  require(tidycensus)
  require(data.table)
  
  state_acs <- 
    get_acs(geography = geography, survey = "acs5", variables = variables, year = year,
            state = state, county = county, geometry = geometry, ...)
  
  # Save immediately after downloading data in case some states fail
  fn <- sprintf("data/acs/%s/%s_%s.rds", geography, state, year)
  
  saveRDS(state_acs, fn)
  
  # return(state_tracts)
  rm(state_acs)
  gc()
}


## CENSUS summary file 3

sf3_state <- function(state, geography, variables, year = 2000, ...) {
  require(tidycensus)
  require(data.table)
  
  state_census <- 
    get_decennial(geography = geography, survey = "sf3", variables = variables, year = year,
                  state = state, ...)
  
  # Save immediately after downloading data in case some states fail
  fn <- sprintf("data/census/%s_%s.rds", state, year)
  saveRDS(state_census, fn)
  
  # return(state_census)
}

```

# ACS data
4x speed up from running in parallel. For some reason, I can't grab the shapefiles at the same time.
```{r tract_par, eval = TRUE}
system.time(
  # For each year
  map(2010:2015,
      # download data for all states
      ~ mclapply(uniq_state,
                 function(st) acs_state(state = st, year = .x, geography = "tract",
                                        variables = vars, geometry = FALSE),
                 mc.cores = 10
      ),
  )
)

```


```{r tract17, eval = TRUE}
# Download computer variables from 2017 (first year available) 
# with the shapefile attached

# No data on territories
lapply(head(uniq_state, -6), 
         function(.x) acs_state(state = .x, geography = "tract", year = 2017, 
                                variables = compvars, geometry = TRUE, 
                                output = "wide")) %>% 
  system.time()
```


```{r}
tract_fn <- list.files("data/acs/tract", pattern = "^[A-Z]{2}_", full = TRUE)

tract_ldf <- mclapply(tract_fn, readRDS)

names(tract_ldf) <- str_extract(tract_fn, "[A-Z_0-9]*(?=\\.rds)")

state_acs10_15 <- rbindlist(tract_ldf[grep("2017", names(tract_ldf), invert = TRUE)], 
                            idcol = "tag", fill = TRUE)

state_acs10_15[ , `:=`(year = str_extract(tag, "\\d{4}"),
                       state = str_extract(tag, "[A-Z]{2}"),
                       tag  = NULL)]

saveRDS(state_acs10_15, "data/acs/alltracts10-15_lng.rds")
```


Don't know why, but some states had tracts saved as POLYGONS. Need to coerce to MULTIPOLYGONS before binding into a single `sf`.
```{r}
ldf17 <- tract_ldf[grep("2017", names(tract_ldf))]
rm(tract_ldf)

poly_ind <- 
  map(ldf17, "geometry") %>% 
  map(class) %>% map( ~ "sfc_POLYGON" %in% .x) %>%
  unlist() %>% which()

ldf17[poly_ind] <- map(ldf17[poly_ind], sf::st_cast, to = "MULTIPOLYGON")

state_acs17 <- rbindlist(ldf17, idcol = "tag", fill = TRUE)

state_acs17[ , `:=`(year = str_extract(tag, "\\d{4}"),
                    state = str_extract(tag, "[A-Z]{2}"),
                    tag  = NULL)]

as.data.frame(state_acs17) %>% sf::st_as_sf() %>% 
  saveRDS("data/acs/alltracts17.rds")
```


# Decennial census data

```{r}
# sf1
# male 65+, sf1: P012020, P012021, P012022, P012023, P012024, P012025,
# female 65+, sf1: P012044, P012045, P012046, P012047, P012048, P012049,


census_vars <- 
  c(
    "P001001", # population total
    "P006002", # white alone
    "P006003", # black alone
    
    "P039008", # male 65+
    "P039019", # female 65+
    
    "P037011", # male HS grads
    "P037028", # female HS grads
    
    "P110004", # ability to speak english
    
    "P087002", # total poverty
    "P054001", # aggregate hh income
    
    "P062002", # SS count
    "P071001", # SS income
    "P073001" # public assistance income
  )


```

```{r eval = TRUE}
system.time(
  mclapply(uniq_state,
           function(st) sf3_state(state = st, geography = "tract", year = 2000, 
                                  variables = census_vars),
           mc.cores = 10
  )
)
```

```{r}
census_fn <- list.files("data/census", pattern = "^[A-Z]{2}_", full = TRUE)

census_ldf <- mclapply(census_fn, readRDS)

names(census_ldf) <- str_extract(census_fn, "[A-Z_0-9]*(?=\\.rds)")

census00 <- rbindlist(census_ldf[grep("2017", names(census_ldf), invert = TRUE)], 
                      idcol = "tag", fill = TRUE)

census00[ , `:=`(year = str_extract(tag, "\\d{4}"),
                       state = str_extract(tag, "[A-Z]{2}"),
                       tag  = NULL)]

saveRDS(census00, "data/census00_lng.rds")
```


