---
title: "Preprocess Census data: Reshaping and interpolation"
author: 
  - "Cole Tanigawa-Lau"
date: ""
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: 3
    toc_float: yes
  theme: paper
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r about, echo=FALSE, results = "asis"}
cat(
  sprintf("Updated: %s.", 
          format(Sys.time(), "%b %d, %Y at %H:%M:%S", usetz = TRUE)),
  sprintf("Working directory: `%s`.", getwd()),
  sep = "\n\n"
)
```


This script prepares the Census demographics and joins it with annual service areas by spatial interpolation.


Reads:

  - data/acs/alltracts10-15_lng.rds
  - data/acs/alltracts17.rds
  - data/service_areas/*

Writes:

  - data/sf_interp/

# Setup

```{r packages, warning=FALSE, message=FALSE}
rm(list = ls())

library(dplyr)
library(tidyr)
library(data.table)
library(coler)
library(sf)
```

Variable info:

```{r}
var_dict <- 
  dplyr::tribble(~ variable,  ~ label, ~ descrip,
               
               "B00001_001",  "pop",  "population count",
               
               "B06001_011",  "age6574",  "65-74",
               "B06001_012",  "age7485",  "74-85",
               "B06001_013",  "age86p",  "86+",
               
               "B99163_004",  "english_imp",  "ability to speak english in population age 5+ (imputed)",
               "B10054_013",  "english_lessvw",  "speak english less than very well",
               
               "B02001_002",  "white",  "total white alone",
               "B02001_003",  "black",  "total black alone",
               
               "B06009_003",  "educ_hsgrad",  "total hs grads",
               "B06009_002",  "educ_hsless",  "total less than hs",
               
               "B07101_002",  "nonmovers",  "non-movers",
               
               "B06011_001",  "income_med",  "median income in past 12 months",
               "B19313_001",  "income",  "aggregate income in past 12 months",
               "B19057_001",  "income_publass",  "public assistance income in past 12 months",
               "B19055_001",   "income_ss",  "social security income in past 12 months",
               
               "B16009_001",  "poverty",  "total poverty status in past 12 months",
               "B06012_002",  "poverty100m",  "total below 100% of poverty level",
               "B06012_003",  "poverty100150",  "total between 100 and 149% of poverty level",
               "B10059_001",  "poverty_grand",  "total poverty status in past 12 months of childrearing grandparents",
               
               
               ################# CENSUS ####################
               "P001001", "pop", "population total",
               "P006002", "white", "white alone",
               "P006003", "black", "black alone",
               "P039008", "male65p", "male 65+",
               "P039019", "female65p", "female 65+",
               "P037011", "educ_hsgrad_male", "male HS grads",
               "P037028", "educ_hsgrad_female", "female HS grads",
               "P110004", "english_imp", "ability to speak english",
               "P087002", "poverty", "total poverty",
               "P054001", "income", "aggregate hh income",
               "P071001", "income_ss", "SS income",
               "P073001", "income_publass", "public assistance income",
               "P062002", "indv_ss", "SS count"
               )
```


# Prepare tract data

## ACS variables

```{r reshape 10-15}
tract_lng <- readRDS("data/acs/alltracts10-15_lng.rds") %>% 
  left_join(var_dict, by = "variable")


tract <- 
  dcast(tract_lng[ , .(year, GEOID, label, estimate)], 
        year + GEOID ~ label, value.var = "estimate")

tract[ , seniors := age6574 + age7485 + age86p
       ][ , c("age6574", "age7485", "age86p") := NULL]

```



## 2000 census

ACS 5-year survey (only one with nationwide coverage) didn't come out until 2010/2011. I need to add the 2000 census data and interpolate for years 2001-2010.
```{r preprocess 00}
tract00_lng <- readRDS("data/census00_lng.rds") %>% 
  left_join(var_dict, by = "variable")

tract00 <- 
  dcast(tract00_lng[ , .(year, GEOID, label, value)], 
        year + GEOID ~ label, value.var = "value")

tract00[ , seniors := male65p + female65p
         ][ ,
            "educ_hsgrad" := educ_hsgrad_male + educ_hsgrad_female
            ]

tract00[ , 
         c("male65p", "female65p", "educ_hsgrad_male", "educ_hsgrad_female") :=
           NULL]
```

```{r}
# Preserves column types
dummy <- copy(tract00)
dummy_geo <- dummy$GEOID
dummy[ ] <- NA

dummy[ , GEOID := dummy_geo]

tract_meat <-
  map(2001:2009, ~ mutate(dummy, year = as.character(.x))) %>% 
  rbindlist()
```


```{r}
# Find variables that are contained in both Census and ACS data
tract_bread <- rbind(tract00, tract, fill = TRUE)
complete_vars <- intersect(names(tract00), names(tract))

# Sandwich the interpolation dummy rows
tract_full <- bind_rows(tract00[ , ..complete_vars],
                        tract_meat[ , ..complete_vars],
                        tract[ , ..complete_vars])


tract_full[ , stfips := str_extract(GEOID, "\\d{2}")]
```

Interpolation between 2000 and 2010 using the `baytrends` package.
```{r}
time_interp_vars <- setdiff(complete_vars, c("year", "GEOID"))

setorder(tract_full, GEOID, year)

system.time(
  tract_full[ , 
              
              (time_interp_vars) := lapply(.SD, baytrends::fillMissing),
              
              by = GEOID,
              .SDcols = time_interp_vars
              ]
)

#   user  system elapsed 
# 45.514   0.109  45.351 

#    user  system elapsed 
# 107.247   1.723 108.675 
```


Still have NA values in `tract_wide` where 2010 tracts did not match to a 2000 tract or vice versa. Tracts are added and removed according to population changes. See [documentation here](https://www2.census.gov/geo/pdfs/education/CensusTracts.pdf), particularly pages 8-13.
```{r}
tract_full[is.na(black)]$year %>% unique()
lapply(tract_full, prop_na) %>% bind_rows()

## Example of GEOID missing from 2000 census data
# tract00[GEOID == "01001020801"]

## Example of GEOID missing after 2000 census
# tract[GEOID == "04001940100"]
```

Pivot to wide format to match `sf` object.
```{r wide for sf}
tract_wide <- 
  pivot_wider(tract_full, id_cols = GEOID, names_from = year,
              values_from = setdiff(complete_vars, c("year", "GEOID")))

lapply(tract_wide, prop_na) %>% bind_rows()

tract_wide %>% filter(substr(GEOID, 1, 2) == "04")
```

## 2017 computer variables

Merge with 2017 computer variables for every year. This data set contains the `sf` object that I need for spatial interpolation below.

```{r read compvars17}
tract17 <- readRDS("data/acs/alltracts17.rds") %>% 
  dplyr::select(GEOID, state, 
                total_comp = B28001_001E, no_comp = B28001_011E, 
                total_internet = B28002_013E)
```

```{r join to sf}
tract_sf <- left_join(x = tract17, y = tract_wide, by = "GEOID") %>% 
  st_transform(st_crs(3857))
```


# Interpolation

For a given state,
  
  1) Load tract demographics and all years' service areas
  2) Compute areal-weighted interpolation from each year's ACS variables to that year's service area
  3) Produce a data set of demographics at the office-year level with stationary computer variables from 2017



## Spatial interpolation setup

Since `st_interpolate_aw()` lacks an `na.rm` parameter, we need to replace NA values with something. These are extensive variables, so I'll replace with 0. See discussion at https://github.com/r-spatial/sf/issues/830.

During interpolation, `nafill` column should count the tracts in which I've replaced some missing value with 0. Remember to check this against `num_tracts` later.

```{r}
num_cols <- sapply(tract_sf, is.numeric) %>% which() %>% 
  names()

tract_sf$num_tracts <- 1

tract_sf$nafill <- 
  dplyr::select(tract_sf, all_of(num_cols)) %>% st_drop_geometry() %>% 
  {!complete.cases(.)} %>% as.numeric()


setnafill(tract_sf, type = "const", fill = 0, cols = num_cols)
```


```{r list setup}
tract_sfl <- split(tract_sf, tract_sf$state)

serv_sfl <- 
  list.files("data/service_areas", full = TRUE) %>% 
  lapply(readRDS)

names(serv_sfl) <- 
  map(serv_sfl, "state") %>% 
  sapply(unique)

interp_fn <- "data/sf_interp/%s.rds"
tract_sfl_select <- 
  map(tract_sfl, ~ dplyr::select(.x, matches("_20|comp|internet")))
```


<!-- Don't evaluate this chunk when knitting. It takes a very long time.  -->
<!-- For some reason, this fails on the polisci server, so I ran it on KBLY. -->

## Run spatial interpolation

```{r interpolation, eval = TRUE}
system.time(
  parallel::mclapply(names(tract_sfl_select), 
                     function(st){
                       
                       # Split service areas by year
                       interp_yrl <- split(serv_sfl[[st]], serv_sfl[[st]]$year)
                       
                       # Run interpolation separately for each year
                       for(yr in as.character(2000:2015)){
                         
                         # Subset to this year's service area
                         sa_sf <- filter(serv_sfl[[st]], year == yr)
                         
                         # Subset to this year's demographics and rename columns
                         demog_sf <- 
                           dplyr::select(tract_sfl_select[[st]], matches(yr)) %>% 
                           
                           rename_with(~ gsub(paste0("_", yr), "", .x))
                         
                         # Interpolate
                         interp <- st_interpolate_aw(demog_sf, to = sa_sf, extensive = TRUE)
                         
                         # Re-link with service areas
                         # Group.1 refers to the row number in `to` (above) 
                         # to which the interpolation corresponds. This fixes
                         # errors in AZ and FL.
                         interp2 <- 
                           sa_sf[interp$Group.1, ] %>% 
                           st_drop_geometry() %>% 
                           bind_cols(interp)
                         
                         # Store in output list
                         interp_yrl[[yr]] <- interp2
                         
                       }
                       
                       
                       out <- bind_rows(interp_yrl)
                       # return(out)
                       saveRDS(out, sprintf(interp_fn, st))
                       
                     }, mc.cores = 5, mc.preschedule = FALSE
  )
)

# user  system elapsed 
# 793.166  36.290 197.002 
#    user  system elapsed 
# 832.362  62.149 207.942 
```



# Export
```{r export}
interp_files <- list.files("data/sf_interp", full = TRUE)
interp <- lapply(interp_files, readRDS) %>% bind_rows()

interp <- dplyr::select(interp, state, year, office, everything(), -geometry)

saveRDS(interp, "data/demograph_service.rds")
```


