# Final data prep
```{r}
rm(list = ls())
```


```{r sub setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 7, fig.height = 5)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

Cutoff function.
```{r}
cutoff <- 
  function(x, threshold = 0.5, levels = 0:1) 
    factor(as.numeric(as.character(x) > threshold), levels = levels)

```

Load data.
```{r}
demog <- readRDS("data/demograph_service.rds") %>% as.data.table() 
closure <- readRDS("data/office_closures.rds") %>% as.data.table()
```

Merge.
```{r}
offices <- merge(closure, demog, all = TRUE, by = c("office", "year"))
```

Subset to complete observations. This removes office-years that have no matching service area, because these offices were closed prior to the the last year of data.
```{r}
offcompl <- offices[complete.cases(offices)]

offcompl[ , .(closures = sum(closed), 
              still_open = .N - sum(closed)),
          by = year]
```



Change outcome to factor for classification trees.
```{r}
offcompl$closed <- as.factor(offcompl$closed)
```

Remove 2015 data since the closure data seem to run through 2014, despite the SSA stating the data come from 2016.

```{r}
offcompl <- offcompl[year %in% as.character(2010:2014)]
```

Test with two data sets: per capita and absolute values. I redefine here.

```{r}
absolute <- offcompl

vars_absolute <- 
  c("white", "black", "english_imp", "income", "income_publass",
    "income_ss", "poverty", "seniors", "educ_hsgrad", 
    "pop",
    
    "total_comp17", "no_comp17", "total_internet17",
    
    "area", "nn_dist", "num_tracts"
    )
```

Check for collinearity between Social Security and public assistance income. Drop if correlation exceeds 0.95.

```{r}
if(offcompl[ , cor(income_ss, income_publass) > 0.95])
  
  vars_absolute <- setdiff(vars_absolute, "income_publass")

form_absolute <- reformulate(termlabels = vars_absolute, response = "closed")
```


```{r}
percap <- copy(offcompl)


# Variables not to divide by population
vars_nopc <- c("pop", "nn_dist", "prop_area", "prop_num_tracts")


set_cols(percap, j = setdiff(vars_absolute, vars_nopc), 
         FUN = function(x) x / percap[["pop"]])

setnames(percap, 
         setdiff(vars_absolute, vars_nopc), 
         paste0("prop_", setdiff(vars_absolute, vars_nopc)))
```


```{r}
vars_percap <- paste0("prop_", setdiff(vars_absolute, vars_nopc)) %>%
  c("pop")

form_percap <- reformulate(termlabels = vars_percap, response = "closed")


```

Some places actually had zero population (interpolated from census tracts).
```{r}
percap <- percap[pop > 0]
```
